---
title: "Testing in R"
output: html_document
date: "2023-06-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mystery Coffee

```{r}
make_groups <- function(names){
  names_shuffled <- sample(names)
  names_coupled <- matrix(names_shuffled, ncol = 2)
  return(names_coupled)
}
```


Get some names:
```{r}
friends <- sample(starwarsdb::people$name, 20)

friends
```
Group them:
```{r}
make_groups(friends)
```

## What do you look for?



## Let's make this explicit: we will write a test

```{r}
# install.packages("testthat")
library(testthat)
```

```{r}
friends <- c("Jar Jar Binks","Han Solo", "Boba Fett", "Anakin Skywalker", "Leia Organa", "C-3PO")

grouped_friends <- make_groups(friends)

# test with testthat functions
expect_equal(dim(grouped_friends), c(3, 2))
expect_equal(class(grouped_friends), c("matrix", "array"))

# any other ideas?
```


## Let's update our function

How does it deal with an uneven number of participants?

```{r}
friends <- sample(starwarsdb::people$name, 5)
make_groups(friends)
```

We don't want people to have a double duty. So let's fix this.

```{r}

make_groups <- function(names){
  # Shuffle names, and remove one if there is an uneven number
  if(length(names)%%2 > 0){
    names_shuffled <- sample(names, length(names)-1)
    message("Removed one person due to an uneven number of participants.")
  } else{
    names_shuffled <- sample(names)
  }

  # Arrange the  names
  names_coupled <- matrix(names_shuffled, ncol = 2)

  return(names_coupled)
}

```


## Add tests

What tests can we run?

```{r}
friends <- c("Jar Jar Binks","Han Solo", "Boba Fett", "Anakin Skywalker", "Leia Organa")

grouped_friends <- make_groups(friends)

expect_equal(dim(grouped_friends), c(2, 2))








expect_no_warning(
  grouped_friends <- make_groups(friends)
  )
```

## Test-driven development

First the test, then the code:

```{r}
# data frames are not able to run:
df <- data.frame(1:10, 11:20)

expect_error(make_groups(df))
```


Now we update the code such that the test passes:

```{r}

make_groups <- function(names){
  if(!class(names) %in% c("character","integer")){
    stop("Input must be a vector.")
  }

  # Shuffle names, and remove one if there is an uneven number
  if(length(names)%%2 > 0){
    names_shuffled <- sample(names, length(names)-1)
    message("Removed one person due to an uneven number of participants.")
  } else{
    names_shuffled <- sample(names)
  }

  # Arrange the  names
  names_coupled <- matrix(names_shuffled, ncol = 2)

  return(names_coupled)
}

```

The test again:

```{r}
# data frames are not able to run:
df <- data.frame(1:10, 11:20)

expect_error(make_groups(df), "Input must be a vector")
```


